---
title: "Assignment 8"
author: "Sumaiya Shefa"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    toc_depth: 3
    css: styles.css
  
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, message=FALSE, include=FALSE, echo=FALSE}
### Test and Train Error Formula ###

## Train Error
# train.pred = predict(fit.train.set, train.set)
# mean(train.pred!= train.set$variable)
# 
# ## Test Error
# test.pred = predict(fit.train.set, test.set)
# mean(test.pred != test.set$variable)
```

## Q.5
#### (a) Generate a data set with n = 500 and p = 2
```{r}
x1=runif (500) -0.5
x2=runif (500) -0.5
y=1*(x1^2-x2^2 > 0)
```

#### (b) Plot the observations, colored according to their class labels.
```{r}
library(ggplot2)

ggplot() + geom_point(
  mapping = aes(x1[y == 0], x2[y == 0]),
  col = "blue",
  alpha = 0.5,
  size = 2
) + geom_point(
  mapping = aes(x1[y == 1], x2[y == 1]),
  col = "orange",
  alpha = 0.8,
  size = 2
) + labs(x = "x1", y = "x2")
```
```{r}
```

#### (c) Fit a logistic regression model
```{r}
df=as.data.frame(cbind(x1, x2, y))

glm.fit = glm(y ~ ., df, family = "binomial")
summary(glm.fit)
```
#### (d) Apply this model to the training data in order to obtain a predicted class label for each training observation. Plot the observations, colored according to the predicted class labels. The decision boundary should be linear.
```{r, message=FALSE}
glm.probs = predict(glm.fit, newdata = df, type = "response")
glm.preds = ifelse(glm.probs > 0.5, 1, 0)
pred1 = df[glm.preds == 1, ]
pred2 = df[glm.preds == 0, ]

ggplot() +
  geom_point(
    data = pred1,
    mapping = aes(x1, x2),
    col = "slateblue",
    alpha = 0.8,
    size = 2
  ) +
  geom_point(
    data = pred2,
    mapping = aes(x1, x2),
    col = "skyblue2",
    alpha = 0.8,
    size = 2
  ) 
```
```{r}
```

#### (e) Now fit a logistic regression model to the data using non-linear functions of X1 and X2 as predictors (e.g. X2^1 , X1×X2, log(X2), and so forth).
```{r, warning=FALSE}
glm.fit2 = glm(y ~ I(x1^2) + I(x2^2) + I(x1^3) + I(x2^3) + (x1 * x2), family = "binomial", data=df)
summary(glm.fit2)
```

#### (f) Apply this model to the training data in order to obtain a predicted class label for each training observation. Plot the observations, colored according to the predicted class labels. The decision boundary should be obviously non-linear. If it is not, then repeat (a)-(e) until you come up with an example in which the predicted class labels are obviously non-linear.

```{r, message=FALSE}
glm.probs2 = predict(glm.fit2, newdata = df, type = "response")
glm.preds2 = ifelse(glm.probs2 > 0.5, 1, 0)

pred3 = df[glm.preds2==1,]
pred4 = df[glm.preds2==0,]

ggplot()+
    geom_point(data=pred3, mapping = aes(x1, x2), col = 
                 "yellowgreen", alpha=0.8, size=2)+
    geom_point(data=pred4, mapping = aes(x1, x2), col = 
                 "turquoise4", alpha=0.5, size=2)

```
```{r}
```
#### (g) Fit a support vector classifier to the data with X1 and X2 as predictors. Obtain a class prediction for each training observation. Plot the observations, colored according to the predicted class labels.
```{r}
library(e1071)

svm.fit <- svm(as.factor(y)~ ., data=df, kernel ="linear", cost=0.1)

summary(svm.fit)
```

```{r}
svm.probs = predict(svm.fit, df)

pred5 = df[svm.probs==1,]
pred6 = df[svm.probs==0,]

ggplot() +
  geom_point(
    data = pred5,
    mapping = aes(x1, x2),
    col = "skyblue" ,
    size = 2, alpha=0.8
  ) +
  geom_point(
    data = pred6,
    mapping = aes(x1, x2),
    col = "orange2",
    size = 2, alpha=0.6
  ) 
plot(svm.fit, df)
```
```{r}
```

#### (h) Fit a SVM using a non-linear kernel to the data. Obtain a class prediction for each training observation. Plot the observations,colored according to the predicted class labels.
```{r}
svm.fit2 <- svm(as.factor(y)~ x1 + x2, data=df, method = "radial", cost=0.1)
summary(svm.fit2)
```

```{r}
svm.probs2 = predict(svm.fit2, df)

pred7 = df[svm.probs2==1,]
pred8 = df[svm.probs2==0,]

ggplot() +
  geom_point(
    data = pred7,
    mapping = aes(x1, x2),
    col = "skyblue" ,
    size = 2, alpha=0.8
  ) +
  geom_point(
    data = pred8,
    mapping = aes(x1, x2),
    col = "orange2",
    size = 2, alpha=0.6
  ) 
plot(svm.fit2, df)
```
```{r}
```

## Q.7
#### (a) Create a binary variable that takes on a 1 for cars with gas mileage above the median, and a 0 for cars with gas mileage below the median.
```{r, message=FALSE}
library(ISLR)
data("Auto")

mpg01 = ifelse(Auto$mpg > median(Auto$mpg), 1,0)
Auto01 = data.frame(Auto[-c(1,9)], mpg01)
```
```{r, message=FALSE, include=FALSE}
attach(Auto01)
```

#### (b) Fit a support vector classifier to the data with various values of cost, in order to predict whether a car gets high or low gas mileage. Report the cross-validation errors associated with different values of this parameter. Comment on your results.
```{r}
set.seed(1)
tune.lin=tune(svm,mpg01~.,data=Auto01,kernel ="linear",ranges=list(cost=c(0.001, 0.01, 0.1, 1,5,10,100)))
summary(tune.lin)
```
Cost = 0.01 results in the lowest cross-validation error rate.
#### (c) Now repeat (b), this time using SVMs with radial and polynomial basis kernels, with different values of gamma and degree and cost. Comment on your results.
```{r}
tune.rad <- tune(svm, mpg01 ~ ., data = Auto01, kernel = "radial", ranges = list(cost = c(0.1, 1, 5, 10), gamma = c(0.01, 0.1, 1, 5, 10, 100)))
summary(tune.rad)
```
Cost = 1 results in the lowest cross-validation error rate.
```{r}
tune.poly <- tune(svm, mpg01 ~ ., data = Auto01, kernel = "polynomial", ranges = list(cost = c(0.1, 1, 5, 10), degree = c(2, 3, 4)))
summary(tune.poly)
```
Cost = 10 and degree = 3 results in the lowest cross-validation error rate.

#### (d) Make some plots to back up your assertions in (b) and (c).Hint: In the lab, we used the plot() function for svm objects only in cases with p = 2. When p > 2, you can use the plot() function to create plots displaying pairs of variables at a time. 
```{r}
svm.lin = svm(as.factor(mpg01) ~ ., data = Auto01, kernel = "linear", cost = 0.01)
svm.poly = svm(as.factor(mpg01) ~ ., data = Auto01, kernel = "polynomial", cost = 10, 
    degree = 3)
svm.rad = svm(as.factor(mpg01) ~ ., data = Auto01, kernel = "radial", cost = 1, gamma = 1)
```
```{r}
Auto02 = data.frame(Auto01, Auto[c(1)])

# for slicing
m.acc <- mean(acceleration)
m.wt <- mean(weight)
m.hp <- mean(horsepower)
m.dp <- mean(displacement)
m.mpg <- mean(Auto$mpg)
```
##### Plots
```{r}
# linear
plot(
  svm.lin,
  Auto02,
  mpg ~ acceleration,
  main = "Linear: mpg~acceleration",
  slice = list(displacement = m.dp, horsepower = m.hp)
)
plot(
  svm.lin,
  Auto02,
  mpg ~ displacement,
  main = "Linear: mpg~displacement",
  slice = list(acceleration = m.acc, horsepower = m.hp)
)
plot(
  svm.lin,
  Auto02,
  mpg ~ horsepower,
  main = "Linear: mpg~horsepower",
  slice = list(acceleration = m.acc, displacement = m.dp)
)
plot(
  svm.lin,
  Auto02,
  mpg ~ weight,
  main = "Linear: mpg~weight",
  slice = list(acceleration = m.acc, displacement = m.dp)
)
plot(
  svm.lin,
  Auto02,
  mpg ~ cylinders,
  main = "Linear: mpg~cylinders",
  slice = list(displacement = m.dp, horsepower = m.hp)
)
plot(
  svm.lin,
  Auto02,
  mpg ~ origin,
  main = "Linear: mpg~origin",
  slice = list(displacement = m.dp, horsepower = m.hp)
)
plot(
  svm.lin,
  Auto02,
  mpg ~ year,
  main = "Linear: mpg~year",
  slice = list(displacement = m.dp, horsepower = m.hp)
)
# Radial plots
plot(
  svm.rad,
  Auto02,
  mpg ~ acceleration,
  main = "Radial: mpg~acceleration",
  slice = list(displacement = m.dp, horsepower = m.hp)
)
plot(
  svm.rad,
  Auto02,
  mpg ~ displacement,
  main = "Radial: mpg~displacement",
  slice = list(acceleration = m.acc, horsepower = m.hp)
)
plot(
  svm.rad,
  Auto02,
  mpg ~ horsepower,
  main = "Radial: mpg~horsepower",
  slice = list(acceleration = m.acc, displacement = m.dp)
)
plot(
  svm.rad,
  Auto02,
  mpg ~ weight,
  main = "Radial: mpg~weight",
  slice = list(acceleration = m.acc, displacement = m.dp)
)
plot(
  svm.rad,
  Auto02,
  mpg ~ cylinders,
  main = "Linear: mpg~cylinders",
  slice = list(displacement = m.dp, horsepower = m.hp)
)
plot(
  svm.rad,
  Auto02,
  mpg ~ origin,
  main = "Linear: mpg~origin",
  slice = list(displacement = m.dp, horsepower = m.hp)
)
plot(
  svm.rad,
  Auto02,
  mpg ~ year,
  main = "Linear: mpg~year",
  slice = list(displacement = m.dp, horsepower = m.hp)
)
# Polynomial
plot(
  svm.poly,
  Auto02,
  mpg ~ acceleration,
  main = "Polynomial: mpg~acceleration",
  slice = list(displacement = m.dp, horsepower = m.hp)
)
plot(
  svm.poly,
  Auto02,
  mpg ~ displacement,
  main = "Polynomial: mpg~displacement",
  slice = list(acceleration = m.acc, horsepower = m.hp)
)
plot(
  svm.poly,
  Auto02,
  mpg ~ horsepower,
  main = "Polynomial: mpg~horsepower",
  slice = list(acceleration = m.acc, displacement = m.dp)
)
plot(
  svm.poly,
  Auto02,
  mpg ~ weight,
  main = "Polynomial: mpg~weight",
  slice = list(acceleration = m.acc, displacement = m.dp)
)
plot(
  svm.poly,
  Auto02,
  mpg ~ cylinders,
  main = "Linear: mpg~cylinders",
  slice = list(displacement = m.dp, horsepower = m.hp)
)
plot(
  svm.poly,
  Auto02,
  mpg ~ origin,
  main = "Linear: mpg~origin",
  slice = list(displacement = m.dp, horsepower = m.hp)
)
plot(
  svm.poly,
  Auto02,
  mpg ~ year,
  main = "Linear: mpg~year",
  slice = list(displacement = m.dp, horsepower = m.hp)
)
```

```{r}
```
## Q.8

```{r, message=FALSE}
library(ISLR)
data("OJ")

oj = OJ
```
```{r, message=FALSE, include=FALSE}
attach(oj)
```

#### (a) Create a training set containing a random sample of 800 observations, and a test set containing the remaining observations.
```{r}
inTrain = sample(1070,800)
train = oj[inTrain,]
test = oj[-inTrain,]
```
#### (b) Fit a support vector classifier to the training data using cost=0.01, with Purchase as the response and the other variables as predictors. Use the summary() function to produce summary statistics, and describe the results obtained.
```{r}
svm.lin2 = svm(Purchase ~ ., kernel = "linear", data = train, cost = 0.01)
summary(svm.lin2)
```

#### (c) What are the training and test error rates?
```{r}
test.pred1=predict(svm.lin2, test)
train.pred1=predict(svm.lin2, train)

TrainErr = mean(train.pred1 != train$Purchase)
TrainErr
TestErr = mean(test.pred1 != test$Purchase)
TestErr
```

#### (d) Use the tune() function to select an optimal cost. Consider values in the range 0.01 to 10.
```{r}
set.seed(1)
tune.lin2 = tune(svm, Purchase~., data=train, kernel ="linear", ranges =list(cost=seq(0.01, 10, 0.333)))

op = tune.lin2$best.parameters$cost
op
```

#### (e) Compute the training and test error rates using this new value for cost.
```{r}
svm.lin4 = svm(Purchase ~ ., kernel = "linear", data = train, cost = op)

test.pred2=predict(svm.lin4, test)
train.pred2=predict(svm.lin4, train)

TrainErr = mean(train.pred2 != train$Purchase)
TrainErr
TestErr = mean(test.pred2 != test$Purchase)
TestErr
```

#### (f) Repeat parts (b) through (e) using a support vector machine with a radial kernel. Use the default value for gamma.
```{r}
svm.rad2 = svm(Purchase ~ ., kernel = "radial", data = train, cost = 0.01)

test.pred3=predict(svm.rad2, test)
train.pred3=predict(svm.rad2, train)

TestErr = mean(train.pred3 != train$Purchase)
TestErr
TrainErr = mean(test.pred3 != test$Purchase)
TrainErr
Accuracy = 1 - TrainErr
Accuracy
```
```{r}
tune.lin3 = tune(svm, Purchase~., data=train, kernel ="linear", ranges =list(cost=seq(0.01, 10, 0.333)))

op = tune.lin3$best.parameters$cost

svm.lin5 = svm(Purchase ~ ., kernel = "radial", data = train, cost = op)

train.pred4=predict(svm.lin5, train)
test.pred4=predict(svm.lin5, test)

TestErr = mean(test.pred4 != test$Purchase)
TestErr
TrainErr = mean(train.pred4 != train$Purchase)
TrainErr
```
#### (g) Repeat parts (b) through (e) using a support vector machine with a polynomial kernel. Set degree=2.
```{r}
svm.poly2 = svm(Purchase ~ ., kernel = "polynomial", data = train, cost = 0.01, degree=2)

train.pred5=predict(svm.poly2, train)
test.pred5=predict(svm.poly2, test)

TestErr = mean(test.pred5 != test$Purchase)
TestErr
TrainErr = mean(train.pred5 != train$Purchase)
TrainErr
```
```{r}
tune.lin4 = tune(svm, Purchase~., data=train, kernel ="polynomial", ranges =list(cost=seq(0.01, 10, 0.333)), degree=2)

op = tune.lin4$best.parameters$cost

svm.lin6 = svm(Purchase ~ ., kernel = "polynomial", data = train, cost = op, degree=2)

train.pred6=predict(svm.lin6, train)
test.pred6=predict(svm.lin6, test)

TestErr = mean(train.pred6 != train$Purchase)
TestErr
TrainErr = mean(test.pred6 != test$Purchase)
TrainErr
Accuracy = 1 - TrainErr
Accuracy
```
#### (h) Overall, which approach seems to give the best results on this data?
```{r}
# Radial Model
TestErr = mean(test.pred4 != test$Purchase)
TestErr
TrainErr = mean(train.pred4 != train$Purchase)
TrainErr
```

The support vector model with radial kernel gives the best results because it has the lowest number of errors and a high accuracy.